Endogenous Heterogeneous Innovation

### 2018-03-16 Paper/Paper.tex

This paper investigates the interpemporal productivity choice that a firm has to make when deciding whether to make incremental or abrupt product innovations. As firms pick innovation paths, they can either innovate their products incrementally, though there is naturally an exhaustion of incremental possibilities that can be added, or abruptly. In the latter case, innovation brings a radical change to production, which take a productivity toll. However, the invention of a new product (or a new generation of the product) opens a whole horizon for new incremental innovations. Hence, firms have different R\&D incentives: right after a radical innovation, firms want to benefit from all the possible new incremental innovation steps that they can make. As they exhaust the realm of incremental ideas, firms turn to abrupt innovation to restart the cycle.

This firm behavior prediction has important aggregate implications. At any given point, some share of firms will be trading current productivity for future one as they choose to engage in abrupt R\&D. However, this productivity decline, which should impact the aggregate measure of TFP in the economy, is not related with sluggish efficiency. It is simply the result of an arbitrage decision that would, ideally, pay off in the future.

This paper is linked to the literature on endogenous growth, where creative destruction by firms is the main driver of growth. In this sense, firms engage in R\&D both to increase their mark-up and to steal the products of other firms. Although the recent literature on this subject has taken into account the idea of R\&D heterogeneity (Akcigit and Kerr, 2016) and diminishing returns of incremental innovation (Acemoglu, Akcigit and Celik, 2014), innovation is always good for productivity, even in the short-term. Moreover, since the focus of Acemoglu et al. (2014) was in the relationship between management and R\&D, they do not endogenize the innovation decision. Additionally, to our knowledge, there is no model that accounts for the possibility of firms innovating abruptly internally, which ignores the temporal dimension of the decision to innovate.

On the empirical side, this paper joins the effort of using patent and patent citation data to discipline productivity in an endogenous framework (e.g. Griliches, 1990). The idea behind this identification is that a patent with more citations has a bigger impact on productivity. Using the rich citation distribution available from the empirical work of Hall, Jaffe and Trajtenberg (2001), we discipline our parameters by imposing patent and citation patterns and quantify the impact of each possible innovation path.

>>>>

This paper investigates the interpemporal productivity choice that a firm has to make when deciding whether to make incremental or abrupt product innovations. Technological innovation has long been viewed as the main driver of economic growth. It is clear that innovativeness comes in many different types, including in their scale, scope, and type of impact. Recent literature on endogenous growth has included heterogeneity in innovation. In this paper, a model is presented in which firms endogenously choose between two different types of innovation: abrupt (major) and incremental (minor). This feature allows the model to explain a larger range of behavior, and has a number of interesting predictions. In this work, we take a first step toward applying the model; a few key parameters are estimated using data on patent citation distributions.

Our analysis is linked to the literature on endogenous growth, where creative destruction by firms is the main driver of growth. In this sense, firms engage in R\&D both to increase their mark-up and to steal the products of other firms. Although the recent literature on this subject has taken into account the idea of R\&D heterogeneity (Akcigit and Kerr, 2016) and diminishing returns of incremental innovation (Acemoglu, Akcigit and Celik, 2014), innovation is always good for productivity, even in the short-term. Moreover, since the focus of Acemoglu et al. (2014) was in the relationship between management and R\&D, they do not endogenize the innovation decision. Additionally, to our knowledge, there is no model that accounts for the possibility of firms innovating abruptly internally, which ignores the temporal dimension of the decision to innovate.

Under certain parameter values, our model has interesting predictions regarding intertemporal productivity choices by innovating firms. This interesting behavior arises when abrupt innovations result in a relatively small or even negative productivity change. In this case, right after a radical innovation, firms want to benefit from all the possible new incremental innovation steps that they can make. Eventually they exhaust the realm of incremental ideas, so firms turn to abrupt innovation, even though that may result in a temporary decrease in productivity. This cyclical firm behavior would  have important aggregate implications. At any given point, some share of firms will be trading current productivity for future productivity as they choose to engage in abrupt R\&D. However, this productivity decline, which should impact the aggregate measure of TFP in the economy, is not related with sluggish efficiency. It is simply the result of an arbitrage decision that would, ideally, pay off in the future.

Before applying the model to intertemporal productivity decisions, however, we must first estimate a few key model parameters using steady-state patent citation data. Many previous papers have used patent and patent citation data to discipline productivity in an endogenous framework (e.g. Griliches, 1990). The idea behind this identification is that a patent with more citations has a bigger impact on productivity. Using the rich citation distribution available from the empirical work of Hall, Jaffe and Trajtenberg (2001), we discipline our parameters by imposing patent and citation patterns and quantify the impact of each possible innovation path.

Model parameters are estimated using maximum likelihood estimation and, as a robustness check, the generalized method of moments. Their robustness is also tested to changes in various theoretical assumptions. Our resulting estimates for average arrival rates of both abrupt and incremental innovations are what we would expect given our theoretical model, and they match well with untargeted moments. However, our estimate of another key model parameter is very unexpected. This estimate indicates that incremental innovations are not subject to diminishing returns, which is a key feature of both our model and those from previous papers. If true, this result precludes many of the interesting behavioral predictions of our model, including the intertemporal TFP choice mentioned above. Considering the importance of this result, we devote a large portion of the paper to investigate the theoretical assumptions and empirical features that are driving it.
